---
title: "project2"
author: "Aksel Vincent Berg & Sergi Alonso"
date: "5/30/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


2.2 - Project Realization (Practical Part I)



Task 0. Prepare the data and the variables.

Split the file into training set and test set.

```{r}

training_set <- read.csv("datos_MGTN.csv", nrows=650, header=FALSE)

test_set <- read.csv("datos_MGTN.csv", skip=650, header=FALSE)

```


```{r}

ozone_training <- training_set[,"V2"]
resistor_training <- training_set[,"V3"]
temperature_training <- training_set[,"V4"]
humidity_training <- training_set[,"V5"]

ozone_test <- test_set[,"V2"]
resistor_test <- test_set[,"V3"]
temperature_test <- test_set[,"V4"]
humidity_test <- test_set[,"V5"]

```


Task 1.

```{r}

plot(resistor_training,
     type="l",
     main="Ozone as a function of time",
     xlab="Time",
     ylab="KΩ"
)

plot(ozone_training,
     type="l",
     main="Ozone reference data as a function of time",
     xlab="Time",
     ylab="µgr/m^{3}"
)

```



Task 2.

```{r}

n_training <- length(ozone_training)
n_test <- length(ozone_test)
n_tot <- n_training + n_test


# Find means and standard deviation.

mean_temperature_training <- mean(temperature_training)
sd_temperature_training <- sd(temperature_training)

mean_ozone_training <- mean(ozone_training)
sd_ozone_training <- sd(ozone_training)

mean_resistor_training <- mean(resistor_training)
sd_resistor_training <- sd(resistor_training)

mean_humidity_training <- mean(humidity_training)
sd_humidity_training <- sd(humidity_training)

temp_norm_training_set <- array(dim=n_training)
ozone_norm_training_set <- array(dim=n_training)
resistor_norm_training_set <- array(dim=n_training)
humidity_norm_training_set <- array(dim=n_training)

# Normalize: subtract the mean and devide by the standard deviation.
for (i in 1:n_training) {
    temp_norm_training_set[i] <- (temperature_training[i] - mean_temperature_training) / sd_temperature_training
    
    ozone_norm_training_set[i] <- (ozone_training[i] - mean_ozone_training) / sd_ozone_training
    
    resistor_norm_training_set[i] <- (resistor_training[i] - mean_resistor_training) / sd_resistor_training
    
    humidity_norm_training_set[i] <- (humidity_training[i] - mean_humidity_training) / sd_humidity_training
    
}

temp_norm_test_set <- array(dim=n_test)
ozone_norm_test_set <- array(dim=n_test)
resistor_norm_test_set <- array(dim=n_test)
humidity_norm_test_set <- array(dim=n_test)

for (i in 1:n_test) {
    temp_norm_test_set[i] <- (temperature_test[i] - mean_temperature_training) / sd_temperature_training
    
    ozone_norm_test_set[i] <- (ozone_test[i] - mean_ozone_training) / sd_ozone_training
    
    resistor_norm_test_set[i] <- (resistor_test[i] - mean_resistor_training) / sd_resistor_training
    
    humidity_norm_test_set[i] <- (humidity_test[i] - mean_humidity_training) / sd_humidity_training
}

```


Task 3.

```{r}

plot(ozone_norm_training_set,
     resistor_norm_training_set,
     main="Normalized reference data as a function of normalized ozone sensor data",
     xlab="Normalized ozone sensor data",
     ylab="Normalized reference data"
)

```



Task 4.

```{r}

# Multiple linear regression.
fit_ozone_norm <- lm(ozone_norm_training_set ~ temp_norm_training_set + resistor_norm_training_set + humidity_norm_training_set)

summary(fit_ozone_norm)


# Predict the training data.
ozone_predicted <- predict(fit_ozone_norm)

# Obtain the RMSE.
RMSE <- 0
for (i in 1:n_training) {
    RMSE <- RMSE + (ozone_predicted[i]*sd_ozone_training + mean_ozone_training - ozone_training[i])**2
}
RMSE <- sqrt(1/n_training * RMSE)
print(RMSE)

```

The obtained β coefficients from the multiple linear regression:
- β0: ~ 0.00
- β1: 0.43
- β2: 0.67
- β3: 0.05

The obtained values from the training set:
- R^2: 0.8813
- RMSE: 14.75


Task 5.

```{r}

# Predict the test data.
data_norm_test_set <- data.frame(c(temp_norm_test_set), c(resistor_norm_test_set), c(humidity_norm_test_set))

# The renaming has to take place in order to make the function work.
colnames(data_norm_test_set) <- c("temp_norm_training_set", "resistor_norm_training_set", "humidity_norm_training_set")

test_data_prediction <- predict(fit_ozone_norm, newdata=data_norm_test_set)


# Denormalize the prediction data.
denomarlized_predicted_test <- array(dim=n_test)

for (i in 1:n_test) {
    denomarlized_predicted_test[i] <- test_data_prediction[i]*sd_ozone_training+mean_ozone_training
}


# Obtain the RMSE.
RMSE <- 0
for (i in 1:n_test) {
    RMSE <- RMSE + (denomarlized_predicted_test[i] - ozone_test[i])**2
}
RMSE = sqrt(1/n_test * RMSE)
print(RMSE)


# Obtain the R-squared.
top <- 0
bottom <- 0
mean_ozone_test <- mean(ozone_test)
for (i in 1:n_test) {
    top <- top + (denomarlized_predicted_test[i] - ozone_test[i])**2
    bottom <- bottom + (mean_ozone_test - ozone_test[i])**2
}
r_squared_test_set <- 1 - top/bottom
print(r_squared_test_set)

```

The obtained values from the test set:
- RMSE: 11.777
- R^2: 0.914



Task 6.

- Black line: Actual, normalized data
- Red line: Predicted data
```{r}

# Training predicted vs training reference data.

plot(ozone_norm_training_set,
     type="l",
     lwd=3,
     main="Training predicted vs training reference data",
     ylab="Normalized",
     xlab="Training data"
)
lines(ozone_predicted, col="red")
legend(x="top",
       legend=c("Training reference data", "Training predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)


# Test predicted vs test refence data.

plot(ozone_norm_test_set,
     type="l",
     lwd=3,
     main="Test predicted vs test reference data",
     ylab="Normalized",
     xlab="Test data"
)
lines(test_data_prediction, col="red")
legend(x="top",
       legend=c("Test reference data", "Test predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)


# All data.

all_pred <- array(dim=n_tot)
all_actual <- array(dim=n_tot)

for (i in 1:n_training) {
    all_pred[i] <- ozone_predicted[i]
    all_actual[i] <- ozone_norm_training_set[i]
}
for (i in 1:n_test) {
    all_pred[n_training+i] <- test_data_prediction[i]
    all_actual[n_training+i] <- ozone_norm_test_set[i]
}

plot(all_actual,
     type="l",
     lwd=3,
     main="All predicted vs all reference data",
     ylab="Normalized",
     xlab="All data"
)
lines(all_pred, col="red")
legend(x="top",
       legend=c("All reference data", "All predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)


plot(all_actual,
     all_pred,
     main="All predicted as a function of reference data",
     xlab="Reference data",
     ylab="Predicted data",
)

```




2.2 - Project Realization (Practical Part II)

```{r}

library(e1071)

fit_ozone_norm_svm <- svm(ozone_norm_training_set ~ temp_norm_training_set + resistor_norm_training_set + humidity_norm_training_set)


# Training data.

pred_svm <- predict(fit_ozone_norm_svm)

plot(ozone_norm_training_set,
     type="l",
     lwd=3,
     main="SVM: training predicted vs training reference data",
     ylab="Normalized",
     xlab="Training data"
)
lines(pred_svm, col="red")
legend(x="top",
       legend=c("Training reference data", "Training predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)

denomarlized_predicted_training_svm <- array(dim=n_training)

for (i in 1:n_training) {
    denomarlized_predicted_training_svm[i] <- pred_svm[i]*sd_ozone_training+mean_ozone_training
}


for (i in 1:n_training) {
    RMSE <- RMSE + (denomarlized_predicted_training_svm[i] - ozone_training[i])**2
    top <- top + (denomarlized_predicted_training_svm[i] - ozone_training[i])**2
    bottom <- bottom + (mean_ozone_training - ozone_training[i])**2
}

r_squared_training_svm <- 1 - top/bottom
RMSE_training_svm <- sqrt(1/n_training * RMSE)

print("SVM: TRAINING DATA")
print(r_squared_training_svm)
print(RMSE_training_svm)
print("")



# Test data.

pred_svm_test <- predict(fit_ozone_norm_svm, data_norm_test_set)

plot(ozone_norm_test_set,
     type="l",
     lwd=3,
     main="SVM: test predicted vs test reference data",
     ylab="Normalized",
     xlab="Test data"
)
lines(pred_svm_test, col="red")
legend(x="top",
       legend=c("Test reference data", "Test predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)

denomarlized_predicted_test_svm <- array(dim=n_test)
for (i in 1:n_test) {
    denomarlized_predicted_test_svm[i] <- pred_svm_test[i]*sd_ozone_training+mean_ozone_training
}


RMSE <- 0
for (i in 1:n_test) {
    RMSE <- RMSE + (denomarlized_predicted_test_svm[i] - ozone_test[i])**2
    top <- top + (denomarlized_predicted_test_svm[i] - ozone_test[i])**2
    bottom <- bottom + (mean_ozone_test - ozone_test[i])**2
}

r_squared_test_svm <- 1 - top/bottom
RMSE_test_svm <- sqrt(1/650 * RMSE)

print("SVM: TEST DATA")
print(r_squared_test_svm)
print(RMSE_test_svm)



# All data.

all_pred_svm <- array(dim=n_tot)
all_actual_svm <- array(dim=n_tot)
for (i in 1:n_training) {
    all_pred_svm[i] <- pred_svm[i]
    all_actual_svm[i] <- ozone_norm_training_set[i]
}
for (i in 1:n_test) {
    all_pred_svm[n_training+i] <- pred_svm_test[i]
    all_actual_svm[n_training+i] <- ozone_norm_test_set[i]
}
plot(all_actual_svm,
     type="l",
     lwd=3,
     main="SVM: all predicted vs all reference data",
     ylab="Normalized",
     xlab="All data"
     )
lines(all_pred_svm, col="red")
legend(x="top",
       legend=c("All reference data", "All predicted"),
       col=c("black", "red"),
       lty=1:1,
       cex=1.0,
       box.lty=0
)

```

