---
title: "project2"
author: "Aksel Vincent Berg & Sergi Alonso"
date: "5/30/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Split the file into training set and test set.

```{r}

data_training_set <- read.csv("datos_MGTN.csv", nrows=650, header=FALSE)

data_test_set <- read.csv("datos_MGTN.csv", skip=650, header=FALSE)

```


Task 1: Obtain the mean of the traning set temperature mew_T.

```{r}

ozone_training <- data_training_set[,"V2"]
resistor_training <- data_training_set[,"V3"]
temperature_training <- data_training_set[,"V4"]
humidity_training <- data_training_set[,"V5"]

ozone_test <- data_test_set[,"V2"]
resistor_test <- data_test_set[,"V3"]
temperature_test <- data_test_set[,"V4"]
humidity_test <- data_test_set[,"V5"]

```


Task 2: Obtain the standard deviation (std) of the training set temperature mew_T.

```{r}

mean_temperature_training <- mean(temperature_training)
sd_temperature_training <- sd(temperature_training)

```


Task 3: Normalize all the samples of the training.

```{r}
n_training <- length(ozone_training)
n_test <- length(ozone_test)
n_tot <- n_training + n_test

# Repeat the process for ozone, resistor data and relative humidity.

mean_ozone_training <- mean(ozone_training)
sd_ozone_training <- sd(ozone_training)

mean_resistor_training <- mean(resistor_training)
sd_resistor_training <- sd(resistor_training)

mean_humidity_training <- mean(humidity_training)
sd_humidity_training <- sd(humidity_training)

temp_norm_training_set <- array(dim=n_training)
ozone_norm_training_set <- array(dim=n_training)
resistor_norm_training_set <- array(dim=n_training)
humidity_norm_training_set <- array(dim=n_training)

# Normalize: subtract the mean and devide by the standard deviation.
for (i in 1:n_training) {
    temp_norm_training_set[i] <- (temperature_training[i] - mean_temperature_training) / sd_temperature_training
    
    ozone_norm_training_set[i] <- (ozone_training[i] - mean_ozone_training) / sd_ozone_training
    
    resistor_norm_training_set[i] <- (resistor_training[i] - mean_resistor_training) / sd_resistor_training
    
    humidity_norm_training_set[i] <- (humidity_training[i] - mean_humidity_training) / sd_humidity_training
    
}


temp_norm_test_set <- array(dim=n_test)
ozone_norm_test_set <- array(dim=n_test)
resistor_norm_test_set <- array(dim=n_test)
humidity_norm_test_set <- array(dim=n_test)

for (i in 1:n_test) {
    temp_norm_test_set[i] <- (temperature_test[i] - mean_temperature_training) / sd_temperature_training
    
    ozone_norm_test_set[i] <- (ozone_test[i] - mean_ozone_training) / sd_ozone_training
    
    resistor_norm_test_set[i] <- (resistor_test[i] - mean_resistor_training) / sd_resistor_training
    
    humidity_norm_test_set[i] <- (humidity_test[i] - mean_humidity_training) / sd_humidity_training
}


```




Task 2: Project Realization (Practical Part I)


Task 2.1

```{r}

plot(resistor_training, type="l")

plot(ozone_training, type="l")


```




```{r}

plot(ozone_norm_training_set, resistor_norm_training_set)


```



Task 4.
```{r}

fit_ozone_norm <- lm(ozone_norm_training_set ~ temp_norm_training_set + resistor_norm_training_set + humidity_norm_training_set)

summary(fit_ozone_norm)

ozone_predicted <- predict(fit_ozone_norm)

RMSE <- 0
for (i in 1:n_training) {
    RMSE <- RMSE + (ozone_predicted[i]*sd_ozone_training+mean_ozone_training - ozone_training[i])**2
}

RMSE <- sqrt(1/n_training * RMSE)
print(RMSE)

plot(predict(fit_ozone_norm), type="l")

```

The obtained values from the training set:
- RMSE: 14.75
- R^2: 0.8813



TASK 5

```{r}

data_norm_test_set <- data.frame(c(temp_norm_test_set), c(resistor_norm_test_set), c(humidity_norm_test_set))

# The renaming has to take place in order to make the function work.
colnames(data_norm_test_set) <- c("temp_norm_training_set", "resistor_norm_training_set", "humidity_norm_training_set")

test_data_prediction <- predict(fit_ozone_norm, newdata=data_norm_test_set)
plot(test_data_prediction, type="l")

denomarlized_predicted_test <- array(dim=n_test)

for (i in 1:n_test) {
    denomarlized_predicted_test[i] <- test_data_prediction[i]*sd_ozone_training+mean_ozone_training
}

RMSE <- 0
for (i in 1:n_test) {
    RMSE <- RMSE + (denomarlized_predicted_test[i] - ozone_test[i])**2
}

top <- 0
bottom <- 0
mean_ozone_test <- mean(ozone_test)
for (i in 1:n_test) {
    top <- top + (denomarlized_predicted_test[i] - ozone_test[i])**2
    bottom <- bottom + (mean_ozone_test - ozone_test[i])**2
}

RMSE = sqrt(1/n_test * RMSE)
r_squared_test_set <- 1 - top/bottom

print(RMSE)
print(r_squared_test_set)

```

The obtained values from the test set:
- RMSE: 11.777
- R^2: 0.914



- Black line: Actual, normalized data
- Red line: Predicted data
```{r}

plot(ozone_norm_training_set, type="l", lwd=3)
lines(ozone_predicted, col="red")

plot(ozone_norm_test_set, type="l", lwd=3)
lines(test_data_prediction, col="red")


all_pred <- array(dim=n_tot)
all_actual <- array(dim=n_tot)

for (i in 1:n_training) {
    all_pred[i] <- ozone_predicted[i]
    all_actual[i] <- ozone_norm_training_set[i]
}
for (i in 1:n_test) {
    all_pred[n_training+i] <- test_data_prediction[i]
    all_actual[n_training+i] <- ozone_norm_test_set[i]
}

plot(all_actual, type="l", lwd=3)
lines(all_pred, col="red")

plot(all_actual, all_pred)

```




Task 2.4) Project Realization (Practical Part II)

```{r}
library(e1071)

fit_ozone_norm_svm <- svm(ozone_norm_training_set ~ temp_norm_training_set + resistor_norm_training_set + humidity_norm_training_set)


# Training set
pred_svm <- predict(fit_ozone_norm_svm)
plot(ozone_norm_training_set, type="l", lwd=3)
lines(pred_svm, col="red")

denomarlized_predicted_training_svm <- array(dim=n_training)

for (i in 1:n_training) {
    denomarlized_predicted_training_svm[i] <- pred_svm[i]*sd_ozone_training+mean_ozone_training
}


for (i in 1:n_training) {
    RMSE <- RMSE + (denomarlized_predicted_training_svm[i] - ozone_training[i])**2
    top <- top + (denomarlized_predicted_training_svm[i] - ozone_training[i])**2
    bottom <- bottom + (mean_ozone_training - ozone_training[i])**2
}

r_squared_training_svm <- 1 - top/bottom
RMSE_training_svm <- sqrt(1/n_training * RMSE)

print("SVM TRAINING")
print(r_squared_training_svm)
print(RMSE_training_svm)
print("")



# Test set
myData2 <- data.frame(c(temp_norm_test_set), c(resistor_norm_test_set), c(humidity_norm_test_set))
colnames(myData2) <- c("temp_norm_training_set", "resistor_norm_training_set", "humidity_norm_training_set")

pred_svm_test <- predict(fit_ozone_norm_svm, data_norm_test_set)
plot(ozone_norm_test_set, type="l", lwd=3)
lines(pred_svm_test, col="red")


denomarlized_predicted_test_svm <- array(dim=n_test)
for (i in 1:n_test) {
    denomarlized_predicted_test_svm[i] <- pred_svm_test[i]*sd_ozone_training+mean_ozone_training
}


RMSE <- 0
for (i in 1:n_test) {
    RMSE <- RMSE + (denomarlized_predicted_test_svm[i] - ozone_test[i])**2
    top <- top + (denomarlized_predicted_test_svm[i] - ozone_test[i])**2
    bottom <- bottom + (mean_ozone_test - ozone_test[i])**2
}
r_squared_test_svm <- 1 - top/bottom
RMSE_test_svm <- sqrt(1/650 * RMSE)

print("SVM TEST")
print(r_squared_test_svm)
print(RMSE_test_svm)
print("")




# Everything
all_pred_svm <- array(dim=n_tot)
all_actual_svm <- array(dim=n_tot)
for (i in 1:n_training) {
    all_pred_svm[i] <- pred_svm[i]
    all_actual_svm[i] <- ozone_norm_training_set[i]
}
for (i in 1:n_test) {
    all_pred_svm[n_training+i] <- pred_svm_test[i]
    all_actual_svm[n_training+i] <- ozone_norm_test_set[i]
}
plot(all_actual_svm, type="l", lwd=3)
lines(all_pred_svm, col="red")




```

